MSCA 31006: Final Project - AQI Forecasting

### 2023-08-13 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

acf(pm2_5_monthly_ts, main = "ACF of Weekly Time Series")
pacf(pm2_5_monthly_ts, main = "PACF of Weekly Time Series")

@@

#### Step 6: Split data into train and test

@@

# Split the data into a training and test dataset
train <- window(pm2_5_weekly_ts, start = c(1999,1), end=c(2020,52))
test <- window(pm2_5_weekly_ts, start = c(2021,1), end=c(2023,6))

@@

#### Step 7: Decompose the weekly plot

@@

# Overall, there is no mixed strong trend that normal seasonality model cannot
# capture.

@@

#### Step 6: Benchmark using snaive model

@@

# Forecast the seasonal naïve for each month of 1980
forecast_snaive <- snaive(train, h=110)
# Plot the forecasts for snaive model
plot(forecast_snaive, main = "PM 2.5 AQI Forecast - SNaive",
         xlab = "Week", ylab = "PM 2.5 AQI")
lines(test)

@@

>>>>

acf(pm2_5_weekly_ts, main = "ACF of Weekly Time Series")
pacf(pm2_5_weekly_ts, main = "PACF of Weekly Time Series")

@@

#### Step 6: Split data into train and test - Weekly

@@

# Split the data into a training and test dataset
train <- window(pm2_5_weekly_ts, start = c(1999,1), end=c(2020,52))
test <- window(pm2_5_weekly_ts, start = c(2021,1), end=c(2023,7))

@@

#### Step 7: Decompose the time series plot - Weekly

@@

# Overall, there is no mixed effect that normal seasonality model cannot capture.

@@

#### Step 8: Benchmark using snaive model - Weekly

@@

# Forecast the seasonal naïve for (2021-01 to 2023-07)
forecast_snaive <- snaive(train, h=110)
# Plot the forecasts for snaive model
plot(forecast_snaive, main = "PM 2.5 AQI Forecast - SNaive",
         xlab = "Week", ylab = "PM 2.5 AQI")
lines(test)

@@

#### Step 9: Split data into train and test - Monthly

# Split the data into a training and test dataset
train_monthly <- window(pm2_5_monthly_ts, start = c(1999,1), end=c(2020,12))
test_monthly <- window(pm2_5_monthly_ts, start = c(2021,1), end=c(2023,7))

#### Step 10: Decompose the time series plot - Monthly

# Looking at spectral analysis
periodogram(train_monthly, log = "no", plot=TRUE, 
            ylab = "Periodogram",
            xlab = "Frequency",
            lwd=2, xlim = c(0, 0.2))

# There are two trends: 
#  1) A typical yearly (12 months) seasonal trend
#  2) A trend that is repeating every 8-9 years (100 months)
#  3) A typical half yearly (6 months) seasonal trend

# Overall, there is no mixed effect that normal seasonality model cannot capture.

# Decompose the series
plot(decompose(train_monthly, type="multiplicative"))

# Understand the seasonality and remove it to see the trend
train_monthly_diff <- diff(train_monthly, lag = 12)
autoplot(train_monthly_diff, ylab="Train Seasonal Differencing (Monthly)")

# Let's look if the series is stationary?
kpss.test(train_monthly_diff)
# The reported p-value is 0.1, which is > 0.05, and would suggest that we fail 
# to reject the null hypothesis of level stationarity (conclusion: stationary)

adf.test(train_monthly_diff)
# The reported p-value of 0.01 (<0.05) so we do reject the null hypothesis that 
# the time series is non-stationary (conclusion: stationary)

acf(train_monthly_diff, main = "ACF of Seasonal Differencing Time Series")
pacf(train_monthly_diff, main = "PACF of Seasonal Differencing Time Series")

#### Step 11: Benchmark using snaive model - Monthly

# Forecast the seasonal naïve for (2021-01 to 2023-07)
forecast_snaive_monthly <- snaive(train_monthly, h=31)
# Plot the forecasts for snaive model
plot(forecast_snaive_monthly, main = "PM 2.5 AQI Forecast - SNaive",
         xlab = "Week", ylab = "PM 2.5 AQI")
lines(test_monthly)

# Compare the forecast with the actual test data by calculating MAPE and MSE
# Mean Absolute Percentage Error (MAPE)
mape_snaive <- mean(abs((test_monthly - forecast_snaive_monthly$mean)/test_monthly))
mape_snaive

# Mean Squared Error (MSE)
mse_snaive <- mean((test_monthly - forecast_snaive_monthly$mean)^2)
mse_snaive
